---
title: "Stat425 HW01"
author: "Yu Wu yuw5"
date: "2/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

**1. Problem 1:** The data set prostate from the faraway library, is from a study on 97 men with
who were due to receive a radical prostatectomy. Make a numerical and graphical summary
of the data. Comment on your results.

```{r}
library(faraway)
library(ggplot2)
summary(prostate)

pairs(prostate)
```



**2. Problem 2:** Show that for the SLR model, the coefficient of determination $R^2$ is equal to the square of the correlation coefficient $r_{XY}^2$.


\begin{equation}
\begin{split}

R^2 &= \frac{FSS}{TSS} \\
&= \frac{\Sigma(\hat{y}-\bar{y})^2}{\Sigma(\bar{x}-\bar{x})^2} \\
\\
\hat{y_i} &= \hat{\beta_0}+\hat{\beta_1x_i} \\
\\
\hat{\beta_0} &= \bar{y}-\hat{\beta_1\bar{x}} \\
\\
\hat{\beta_1} &= \frac{S_{XY}}{S_{XX}} \\
&= \frac{\Sigma(x_i-\bar{x})\Sigma(y_i-\bar{y})}{\Sigma(x_i-\bar{x})^2} \\
\\
\Sigma(\hat{y_i}-\bar{y})^2 &= \Sigma(\hat{\beta_0}+\hat{\beta_1}x_i-\bar{y})^2 \\
&= \Sigma{\bar{y}-\hat{\beta_1}\bar{x}+\hat{\beta_1}x_i-\bar{y}} \\
&= \hat{\beta_1}^2\Sigma(x_i-\bar{x})^2 \\
&= \frac{[\Sigma(x_i-\bar{x})\Sigma(y_i-\bar{y})]^2}{[\Sigma(x_i-\bar{x})^2]^2} \Sigma(x_i-\bar{x})^2 \\
&= \frac{[\Sigma(x_i-\bar{x})\Sigma(y_i-\bar{y})]^2}{\Sigma(x_i-\bar{x})^2} \\
\\
R^2 &= \frac{\Sigma(\hat{y}-\bar{y})^2}{\Sigma(\bar{x}-\bar{x})^2} \\
&= \frac{[\Sigma(x_i-\bar{x})\Sigma(y_i-\bar{y})]^2}{\Sigma(x_i-\bar{x})^2 \Sigma(y_i-\bar{y})^2} \\
&= \frac{S_{XY}^2}{S_{XX}S_{YY}} \\
&= r_{XY}^2

\end{split}
\end{equation}



**3. Problem 3:** The manager of the purchasing department of a large company would like to
develop a regression model to predict the average amount of time it takes to process a given
number of invoices. Over a 30-day period, data are collected on the number of invoices
processed and the total time taken (in hours). The data are available in the file invoices.txt.
The following model was fit to the data: Y = β0 + β1x + e, where Y is the processing time
and x is the number of invoices.

(a) Plot the data and comment on the results.   

```{r}
invoice <- read.table('invoices.txt', header = TRUE)

ggplot(invoice, aes(Invoices,Time)) + geom_point(shape=1) + 
  geom_smooth(method=lm, se=FALSE, lwd=1.2)
```

The ls model fits the dataset well. 

(b) Find a 95% confidence interval for the start-up time, ie., β0.  

```{r}
out <- lm(Time~Invoices, data = invoice)
summary(out)
predict(out, newdata=data.frame(Invoices = 0), interval="confidence")
```

(c) Suppose that a best practice benchmark for the average processing time for an additional
invoice is 0.01 hours (pr. 0.6 minutes). Test the null hypothesis H0 : β1 = 0.01 against
a two-sided alternative. Interpret your result.   

```{r}
confint(out, 'Invoices', levels=0.95)
```

The value of $\beta_1$ of $H_0$ is in the confidence interval thus we do not reject the null hypothesis.

(d) Find a point estimate and a 95% prediction interval for the time taken to process 130
invoices.   

```{r}
predict(out, newdata=data.frame(Invoices = 130), interval="prediction")
```



**5. Problem 5:** A story by James R. Hagerty entitled With Buyers Sidelined, Home Prices Slide
published in the Thursday October 25, 2007 edition of the Wall Street Journal contained data
on so-called fundamental housing indicators in major real estate markets across the US. The
author argues that. . . prices are generally falling and overdue loan payments are piling up.
Thus, we shall consider data presented in the article on:
Y = Percentage change in average price from July 2006 to July 2007 (based on the S& P/CaseShiller national housing index); and
x= Percentage of mortgage loans 30 days or more overdue in latest quarter (based on data
from Equifax and Moody’s).
The data are available in the file indicators.txt. Fit the following model to the data: Y =
β0 + β1x + e. Complete the following tasks:

(a) Calculate the R2 and adjusted R2
for the SLR model. Provide an interpretation of both
quantities.

```{r}
indicators <- read.table('indicators.txt', header = TRUE)
out <- lm(PriceChange~LoanPaymentsOverdue, data = indicators)
summary(out)
summary(out)$r.sq
summary(out)$adj.r

```

(b) Find a 95% confidence interval for the slope of the regression model, β1 . On the basis of
this confidence interval decide whether there is evidence of a significant negative linear
association.

```{r}
confint(out, 'LoanPaymentsOverdue', level = 0.95)
```

(c) Use the fitted regression model to estimate E(Y |X = 4). Find a 95% confidence interval
for E(Y |X = 4). Is 0% a feasible value for E(Y |X = 4)? Give a reason to support your
answer.

```{r}
predict(out, data.frame(LoanPaymentsOverdue = 4), interval="confidence")
```

0% is not a feasible value because it is not in the confidence interval. 




















